---
title: "Clustering Cost and Production by ADP"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(DBI)
library(pool)
```


There are three data sources to do this:
1. Cost (Draft data)
2. Production (From Armchair Analysis DB)
3. ADP data (scraped from the web)

Draft data
```{r}
load(file = "/home/john/projects/fantasy-football/data/bn-draft-data/draft_data.Rda")
```

Production Data
```{r}
pool <- dbPool(drv = RMySQL::MySQL(), 
               dbname = "armchair_analysis", 
               host = "localhost", 
               username = "john", 
               port = 3306, 
               unix.sock = "/var/run/mysqld/mysqld.sock")
team_name_abbr_lookup <- data.frame(abbr = c('ARI','ATL','BAL','BUF','CAR','CHI','CIN','CLE','DAL','DEN',
                                             'DET','GB' ,'HOU','IND','JAC','KC' ,'LA' ,'MIA','MIN','NE',
                                             'NO' ,'NYG','NYJ','OAK','PHI','PIT','SD' ,'SEA','SF' , 'STL',
                                             'TB','TEN','WAS'),
                                    name = c('Cardinals','Falcons','Ravens','Bills','Panthers','Bears','Bengals','Browns',
                                             'Cowboys','Broncos','Lions','Packers','Texans','Colts','Jaguars','Chiefs','Rams',
                                             'Dolphins','Vikings','Patriots','Saints','Giants','Jets','Raiders','Eagles',
                                             'Steelers','Chargers','Seahawks','49ers','Rams','Buccaneers','Titans','Redskins'),
                                    stringsAsFactors = FALSE)

offense <- dbGetQuery(con = pool,
                      "SELECT 
                      a.player
                      , b.wk as week
                      , a.year
                      , a.team
                      , 0.04*py + -2*ints + 4*tdp + 0.1*ry + 0.1*recy + 6*tdr + 6*tdrec -2*fuml +6*tdret as fp
                      ,py
                      ,ints
                      ,tdp
                      ,ry
                      ,recy
                      ,tdr
                      ,tdrec
                      ,fuml
                      ,tdret
                      FROM offense as a
                      JOIN game as b on a.gid = b.gid") 


two_point <- dbGetQuery(con = pool,
                        "select 
                        a.*
                        , b.gid
                        , c.seas as year
                        , c.wk as week 
                        from conv as a 
                        join play as b on a.pid = b.pid
                        join game as c on b.gid = c.gid where conv = 1")


rush_two_point <- two_point %>%
  filter(bc != "")
pass_two_point <- two_point %>%
  filter(psr != "")
two_point %>%
  filter(bc == "", psr == "") # forget this one


offense2 <-
  offense %>%
  left_join(., 
            pass_two_point %>% group_by(week,year,psr) %>% summarise(count = n()), 
            by = c("year","week","player" = "psr")) %>% 
  mutate(fp = ifelse(is.na(count) == FALSE, fp + 2*count, fp)) %>%
  select(-count)

offense3 <-
  offense2 %>%
  left_join(.,
            pass_two_point %>% group_by(week,year,trg) %>% summarise(count= n()),
            by = c("year","week","player" = "trg")) %>%
  mutate(fp = ifelse(is.na(count) == FALSE, fp+2*count, fp)) %>%
  select(-count)

offense4 <-
  offense3 %>%
  left_join(.,
            rush_two_point %>% group_by(week,year,bc) %>% summarise(count = n()),
            by = c("year","week","player" = "bc")) %>%
  mutate(fp = ifelse(is.na(count) == FALSE, fp+2*count, fp)) %>%
  select(-count)

offense5 <-
  offense4 %>%
  filter(week <= 16) %>%
  group_by(player,year) %>%
  summarise(ppg= mean(fp),
            ppg_sd = sd(fp),
            gp = n()) %>%
  filter(gp >= 4)


positions <- dbGetQuery(con = pool,
                  "SELECT 
                   a.player
                  ,a.fname
                  ,a.lname
                  ,a.dob
                  ,a.start
                  ,a.pos1 as position
                  FROM player AS a")

offense6 <- offense5 %>%
  left_join(positions, by = "player")


```



ADP Data
```{r}
load("/home/john/projects/fantasy-football/data/adp-data/adp.Rda")
```



Need a Dataset with
|player|year|pos|pos_adp|cost|fp|

```{r}
cost_adp_prod <-
offense6 %>%
  ungroup() %>%
  select(player,year,position,ppg,ppg_sd) %>%
  inner_join(adp %>% 
               ungroup() %>%
               select(year,player_code,pos_adp),
             by = c("player" = "player_code","year")) %>%
  left_join(draft_data %>% select(year,player_code,player,keeper,adj_value),
             by = c("player" = "player_code","year")) 

save(cost_adp_prod,file="/home/john/projects/fantasy-football/data/clustering-data/cost_adp_prod.Rda")
```



```{r}
cost_adp_prod %>%
  filter(position == 'QB', keeper == 0,year==2019) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```



```{r}
cost_adp_prod %>%
  filter(position == 'RB', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```


```{r}
cost_adp_prod %>%
  filter(position == 'WR', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```

```{r}
cost_adp_prod %>%
  filter(position == 'TE', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```

```{r}
table(cost_adp_prod$position)
```


```{r}
sum(is.na(cost_adp_prod$adj_value))
```

```{r}
sum(is.na(cost_adp_prod$keeper))
```



What is the first year of adjusted values?

```{r}
with(cost_adp_prod,plot(year,adj_value))
```

```{r}
table(adp$year)
table(draft_data$year)
table(offense6$year)
```


```{r}
cost_adp_prod_cleaned <- 
  cost_adp_prod %>% 
  filter(keeper == 0 | is.na(keeper), 
         (is.na(adj_value) | adj_value < 200),
         year >= 2010) %>%
  mutate(adj_value = ifelse(is.na(adj_value),0,adj_value))
save(cost_adp_prod_cleaned,file="/home/john/projects/fantasy-football/data/clustering-data/cost_adp_prod_cleaned.Rda")
```

```{r}
head(cost_adp_prod_cleaned)
```

```{r}
sum(cost_adp_prod_cleaned$adj_value == 0)
```



# scale variables by position
```{r}
qb_data <- cost_adp_prod_cleaned %>% filter(position == 'QB')
qb_to_scale <- qb_data[,c(4,5,9)]
qb_to_scale_scaled <- as.data.frame(scale(qb_to_scale))
colnames(qb_to_scale_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(qb_to_scale_scaled)
var(qb_to_scale_scaled[,3])
qb_data_scaled <- bind_cols(qb_data,qb_to_scale_scaled)
```

```{r}
rb_data <- cost_adp_prod_cleaned %>% filter(position == 'RB')
rb_to_scale <- rb_data[,c(4,5,9)]
rb_to_scale_scaled <- as.data.frame(scale(rb_to_scale))
colnames(rb_to_scale_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(rb_to_scale_scaled)
var(rb_to_scale_scaled[,3])
rb_data_scaled <- bind_cols(rb_data,rb_to_scale_scaled)
```

```{r}
wr_data <- cost_adp_prod_cleaned %>% filter(position == 'WR')
wr_to_scale <- wr_data[,c(4,5,9)]
wr_to_scale_scaled <- as.data.frame(scale(wr_to_scale))
colnames(wr_to_scale_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(wr_to_scale_scaled)
var(wr_to_scale_scaled[,1])
wr_data_scaled <- bind_cols(wr_data,wr_to_scale_scaled)
```

```{r}
te_data <- cost_adp_prod_cleaned %>% filter(position == 'TE')
te_to_scale <- te_data[,c(4,5,9)]
te_to_scale_scaled <- as.data.frame(scale(te_to_scale))
colnames(te_to_scale_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(te_to_scale_scaled)
var(te_to_scale_scaled[,3])
te_data_scaled <- bind_cols(te_data,te_to_scale_scaled)
```


# put it back together
```{r}
cost_adp_prod_cleaned_scaled <- bind_rows(qb_data_scaled,
                                  rb_data_scaled,
                                  wr_data_scaled,
                                  te_data_scaled)
save(cost_adp_prod_cleaned_scaled,file="/home/john/projects/fantasy-football/data/clustering-data/cost_adp_prod_cleaned_scaled.Rda")
```



## Plots of Production vs. Cost
 
```{r}
ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'QB'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'RB'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'WR'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'TE'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()
```


```{r}
ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'QB'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'RB'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'WR'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'TE'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()
```



## Now adj_value by pos_adp
```{r}
ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'QB'),
       aes(x = pos_adp, y = adj_value)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'RB'),
       aes(x = pos_adp, y = adj_value)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'WR'),
       aes(x = pos_adp, y = adj_value)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'TE'),
       aes(x = pos_adp, y = adj_value)) +
  geom_point() +
  geom_smooth()
```

Finally, sd by adp

```{r}
ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'QB'),
       aes(x = pos_adp, y = ppg_sd)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'RB'),
       aes(x = pos_adp, y =  ppg_sd)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'WR'),
       aes(x = pos_adp, y =  ppg_sd)) +
  geom_point() +
  geom_smooth()

ggplot(data = cost_adp_prod_cleaned %>% filter(position == 'TE'),
       aes(x = pos_adp, y =  ppg_sd)) +
  geom_point() +
  geom_smooth()
```

