---
title: "WR Projections"
output: html_notebook
---


Here's the idea for WR projections:

Using historical data, build a clustering algorithm for each season. Cluster on total fantasy points scored.

In each season, evaluate variance explained based on cluster size.

Pick one cluster size to apply to all seasons.

Use that cluster size and determine what the average value is for the 1st, 2nd, 3rd, ... cluster.

Build a projection system that analyzes the probability of being in 1st cluster or not, 2nd or not, etc.

Normalize those predictions and then calculate a projected points value as an expected value of cluster and cluster points.

This will be a high bias, low variance system, I think.


```{r}
library(tidyverse)
load("/home/john/projects/fantasy-football/data/analysis-data/wr.Rda")
```


Initial clustering:

Use 2000-2009


```{r}
n_centers <- 10
seasons <- 10
m <- matrix(0,nrow = n_centers*seasons,ncol = 3)
output <- list()
counter <- 1
for(y in c(2000:2009)){
  
  df <- wr %>% filter(year == y, league_rank <= 40)
  
  for(centers in c(1:10)){
  
    k <- kmeans(df$total_points,centers,nstart = 20)
    
    df$cluster <- k$cluster
    
    output[[counter]] <- list()
    output[[counter]] <- df %>%
      group_by(cluster) %>%
      summarise(count = n(),
                mean_points = mean(total_points)) %>%
      arrange(desc(mean_points)) %>%
      mutate(tier = row_number())
    
    m[counter,] <- c(y,centers,k$betweenss/k$totss)
  
    counter <- counter+1
    
    rm(k)
  }
}
  
```



Plot the results:
```{r}
for(y in c(2000:2009)){
  plot <-
  ggplot(data = m %>% as.data.frame() %>% filter(V1 == y),
         aes(x = V2, y = V3)) +
    geom_point() +
    geom_line() +
    ggtitle(paste0("Year ",as.character(y)))
  print(plot)
  rm(plot)
}
```

Five groups seems to work well...

```{r}
ix <- seq(from = 5, to = 5 + n_centers*seasons, by = 10)

output[ix]

five_groups <- do.call(rbind,output[ix])
five_groups$year <- rep(c(2000:2009),each = 5)

ggplot(data = five_groups,
       aes(x = as.factor(tier), y = mean_points)) +
  geom_boxplot() +
  geom_point()
```

```{r}
five_groups %>%
  group_by(tier) %>%
  summarise(min_points = min(mean_points),
            max_points = max(mean_points),
            mean_points = mean(mean_points),
            min_count = min(count),
            max_count = max(count),
            mean_count = mean(count))
```

Look at these values over time...

```{r}
ggplot(data = five_groups,
       aes(x = year, y = mean_points,group = tier, colour = tier)) +
  geom_point() +
  geom_line()

```


```{r}
ggplot(data = five_groups,
       aes(x = year, y = mean_points/16,group = tier, colour = tier)) +
  geom_point() +
  geom_line()

```

