---
title: "WR Projections"
output: html_notebook
---


Here's the idea for WR projections:

Using historical data, build a clustering algorithm for each season. Cluster on total fantasy points scored.

In each season, evaluate variance explained based on cluster size.

Pick one cluster size to apply to all seasons.

Use that cluster size and determine what the average value is for the 1st, 2nd, 3rd, ... cluster.

Build a projection system that analyzes the probability of being in 1st cluster or not, 2nd or not, etc.

Normalize those predictions and then calculate a projected points value as an expected value of cluster and cluster points.

This will be a high bias, low variance system, I think.


```{r}
library(tidyverse)
load("/home/john/projects/fantasy-football/data/analysis-data/wr.Rda")
```


Initial clustering:

Use 2000-2009


```{r}
n_centers <- 10
seasons <- 10
m <- matrix(0,nrow = n_centers*seasons,ncol = 3)
output <- list()
counter <- 1
for(y in c(2000:2009)){
  
  df <- wr %>% filter(year == y, league_rank <= 40)
  
  for(centers in c(1:10)){
  
    k <- kmeans(df$total_points,centers,nstart = 20)
    
    df$cluster <- k$cluster
    
    output[[counter]] <- list()
    output[[counter]] <- df %>%
      group_by(cluster) %>%
      summarise(count = n(),
                mean_points = mean(total_points)) %>%
      arrange(desc(mean_points)) %>%
      mutate(tier = row_number())
    
    m[counter,] <- c(y,centers,k$betweenss/k$totss)
  
    counter <- counter+1
    
    rm(k)
  }
}
  
```



Plot the results:
```{r}
for(y in c(2000:2009)){
  plot <-
  ggplot(data = m %>% as.data.frame() %>% filter(V1 == y),
         aes(x = V2, y = V3)) +
    geom_point() +
    geom_line() +
    ggtitle(paste0("Year ",as.character(y)))
  print(plot)
  rm(plot)
}
```

Five groups seems to work well...

```{r}
ix <- seq(from = 5, to = 5 + n_centers*seasons, by = 10)

output[ix]

five_groups <- do.call(rbind,output[ix])
five_groups$year <- rep(c(2000:2009),each = 5)

ggplot(data = five_groups,
       aes(x = as.factor(tier), y = mean_points)) +
  geom_boxplot() +
  geom_point()
```

```{r}
five_groups %>%
  group_by(tier) %>%
  summarise(min_points = min(mean_points),
            max_points = max(mean_points),
            mean_points = mean(mean_points),
            min_count = min(count),
            max_count = max(count),
            mean_count = mean(count))
```

Look at these values over time...

```{r}
ggplot(data = five_groups,
       aes(x = year, y = mean_points,group = tier, colour = tier)) +
  geom_point() +
  geom_line()

```


```{r}
ggplot(data = five_groups,
       aes(x = year, y = mean_points/16,group = tier, colour = tier)) +
  geom_point() +
  geom_line()

```

## There are two outlier seasons in the data, causing a top tier of size 1. See what happens if we remove those.

```{r}
five_groups %>%
  filter(count == 1)
```

```{r}
wr %>%
  filter(total_points %in% c(265.3,215.0))
```

```{r}
wr_filtered <- wr %>%
  filter(!(total_points %in% c(265.3,215.0)))
```

```{r}
n_centers <- 5
seasons <- 10
m <- matrix(0,nrow = 5*seasons,ncol = 3)
output2 <- list()
counter <- 1

for(y in c(2000:2009)){
  
  df <- wr_filtered %>% filter(year == y, league_rank <= 40)
  
    k <- kmeans(df$total_points,n_centers,nstart = 20)
    
    df$cluster <- k$cluster
    
    output2[[counter]] <- list()
    output2[[counter]] <- df %>%
      group_by(cluster) %>%
      summarise(count = n(),
                mean_points = mean(total_points)) %>%
      arrange(desc(mean_points)) %>%
      mutate(tier = row_number())
    
    m[counter,] <- c(y,n_centers,k$betweenss/k$totss)
    counter <- counter+1
    rm(k)
}
```

```{r}
output2
```



```{r}
five_groups <- do.call(rbind,output2)
five_groups$year <- rep(c(2000:2009),each = 5)
five_groups
```

```{r}
five_groups %>%
  group_by(tier) %>%
  summarise(mean_points = mean(mean_points))
```


```{r}
head(wr)
```



How do the tiers compare to straight league ranks?

```{r}
five_groups %>%
  group_by(tier) %>%
  summarise(mean_points = mean(mean_points),
            mean_count = mean(count))
```

The spread is 20-30 points. 

```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(rank_tier = case_when(league_rank <= 5 ~ 1,
                               between(league_rank,6,10) ~ 2,
                               between(league_rank,11,20) ~ 3,
                               between(league_rank,21,40) ~ 4,
                               TRUE ~ NA_real_)) %>%
  group_by(rank_tier) %>%
  summarise(mean_points = mean(total_points)) %>%
  mutate(mean_ppg = mean_points/16)
```



```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(rank_tier = case_when(league_rank <= 5 ~ 1,
                               between(league_rank,6,10) ~ 2,
                               between(league_rank,11,20) ~ 3,
                               between(league_rank,21,30) ~ 4,
                               between(league_rank,31,40) ~ 5,
                               TRUE ~ NA_real_)) %>%
  group_by(rank_tier) %>%
  summarise(mean_points = mean(total_points)) %>%
  mutate(mean_ppg = mean_points/16) 
```

```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(rank_tier = case_when(between(league_rank,1,8) ~ 1,
                               between(league_rank,9,16) ~ 2,
                               between(league_rank,17,24) ~ 3,
                               between(league_rank,25,32) ~ 4,
                               between(league_rank,33,40) ~ 5,
                               TRUE ~ NA_real_)) %>%
  group_by(rank_tier) %>%
  summarise(mean_points = mean(total_points)) %>%
  mutate(mean_ppg = mean_points/16)
```

```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(rank_tier = case_when(between(league_rank,1,8) ~ 1,
                               between(league_rank,9,16) ~ 2,
                               between(league_rank,17,24) ~ 3,
                               between(league_rank,25,32) ~ 4,
                               between(league_rank,33,40) ~ 5,
                               TRUE ~ NA_real_)) %>%
  group_by(rank_tier) %>%
  summarise(mean_points = mean(total_points)) %>%
  mutate(mean_ppg = mean_points/16)
```

```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(group = case_when(between(league_rank,1,5) ~ 1,
                           between(league_rank,6,10) ~ 1.5,
                           between(league_rank,11,20) ~ 2,
                           between(league_rank,21,30) ~ 3,
                           between(league_rank,31,40) ~ 4,
                           TRUE ~ NA_real_)) %>%
  ggplot(data = .,
         aes(x = league_rank,
             y = total_points)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  geom_smooth(aes(colour = as.factor(group), group = group),method = "lm") +
  geom_smooth(method = 'lm')
```

```{r}
wr %>%
  filter(league_rank <= 40, year > 2010) %>%
  mutate(group = case_when(between(league_rank,1,5) ~ 1,
                           between(league_rank,6,10) ~ 1.5,
                           between(league_rank,11,20) ~ 2,
                           between(league_rank,21,30) ~ 3,
                           between(league_rank,31,40) ~ 4,
                           TRUE ~ NA_real_)) %>%
  ggplot(data = .,
         aes(x = league_rank,
             y = total_points)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  geom_smooth(aes(colour = as.factor(group), group = group),method = "lm") +
  geom_smooth(method = 'lm')
```


```{r}

```




How can we get them the most evenly spaced...

What divides into 40


```{r}
divisors <- c(1,2,4,5,8,10,20,40)
group_size <- 40/divisors
?Hmisc::cut2
```

```{r}
wr %>%
  filter(league_rank <= 40) %>%
  mutate(group = case_when(between(league_rank,))
  
```


## Need two things:
1) P(top 40)
2) P(being in each tier) really 5 things


## Need to build a pipeline for doing this...

Should we filter out the seasons that result in tier 1 being of only size 1 - no I'm thinking now.

What we want to have is 

|               | lag these values        |           | lag these values                                   |
| player | year | team_rank | league_rank | exp | age | tier 1 | tier 2 | tier 3 | tier 4 | tier 5 | top40 |


The alternative would be to do league rank 1-5, 6-10, 11-20, 21-40.

```{r}
wr %>%
  filter(league_rank <= 36, year > 2010) %>%
  mutate(group = case_when(between(league_rank,1,6) ~ 1,
                           between(league_rank,7,12) ~ 2,
                           between(league_rank,13,18) ~ 3,
                           between(league_rank,19,24) ~ 4,
                           between(league_rank,25,30) ~ 5,
                           between(league_rank,31,36) ~ 6,
                           TRUE ~ NA_real_)) %>%
  ggplot(data = .,
         aes(x = league_rank,
             y = total_points)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  geom_smooth(method = 'lm') +
  geom_smooth(aes(colour = as.factor(group), group = group),method = "lm") +
  geom_smooth(aes(colour = as.factor(group), group = group),method = "loess")

```

What are the point drop off values between each?

For example the drop off in average points from 30 to 36 is the same as between 1 and x?

```{r}

```

