---
title: "top40-models"
output: html_document
---

```{r}
library(tidyverse)
load("/home/john/projects/fantasy-football/data/analysis-data/wr.Rda")
```

```{r}
head(wr)
```

```{r}
min(wr$experience)
```

```{r}
wr %>%
  filter(experience == 0) %>%
  mutate(top40 = ifelse(league_rank <= 25,1,0)) %>%
  group_by(year) %>%
  summarise(count = n(),
            sum = sum(top40),
            ttl = sum(top40)/n())
```

```{r}
wr %>%
  filter(league_rank <= 25, year > 2010) %>%
  mutate(group = case_when(between(league_rank,1,5) ~ 1,
                           between(league_rank,6,10) ~ 2,
                           between(league_rank,11,15) ~ 3,
                           between(league_rank,16,20) ~ 4,
                           between(league_rank,21,25) ~ 5,
                           TRUE ~ NA_real_)) %>%
  ggplot(data = .,
         aes(x = league_rank,
             y = total_points)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  geom_smooth(aes(colour = as.factor(group), group = group),method = "lm") +
  geom_smooth(method = 'lm')
```

```{r}
wr %>%
  filter(league_rank <= 25) %>%
  mutate(group = case_when(between(league_rank,1,5) ~ 1,
                           between(league_rank,6,10) ~ 2,
                           between(league_rank,11,15) ~ 3,
                           between(league_rank,16,20) ~ 4,
                           between(league_rank,21,25) ~ 5,
                           TRUE ~ NA_real_)) %>%
  group_by(group) %>%
  summarise(mean_points = mean(total_points))
```

```{r}
wr %>%
  filter(league_rank <= 25) %>%
  mutate(era = ifelse(year >= 2015,1,0)) %>%
  ggplot(data = .,
         aes(x = league_rank,
             y = total_points,
             group=era,
             colour = as.factor(era))) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  geom_smooth(aes(colour = as.factor(era), group = era),method = "lm") +
  geom_smooth(method = 'lm')
```

```{r}
wr %>%
  filter(league_rank <= 25) %>%
  group_by(year) %>%
  summarise(mean_top25_points = mean(total_points)) %>%
  ggplot(data = .,
         aes(x = year,
             y = mean_top25_points)) +
  geom_bar(stat = 'identity')
```

Model baseline:




```{r}
yoy_top <- 
wr %>%
  filter(is.na(league_rank) == FALSE) %>%
  mutate(top25 = ifelse(league_rank <= 25,1,0),
         top10 = ifelse(league_rank <= 10,1,0),
         top5  = ifelse(league_rank <= 5,1,0)) %>%
  arrange(player,year) %>%
  group_by(player) %>%
  mutate(top25_lastyear = ifelse(lag(year) == year - 1, lag(top25),NA),
         top10_lastyear = ifelse(lag(year) == year - 1, lag(top10),NA),
         top5_lastyear  = ifelse(lag(year) == year - 1, lag(top5),NA),
         team_rank_lastyear = ifelse(lag(year) == year - 1,lag(team_rank),NA),
         league_rank_lastyear = ifelse(lag(year) == year - 1,lag(league_rank),NA))
```

```{r}
yoy_top %>%
  filter(top25 == 1) %>%
  group_by(year) %>%
  summarise(total_players = sum(top25),
            carryover25 = sum(top25_lastyear, na.rm = T),
            carryover10 = sum(top10_lastyear, na.rm = T),
            carryover5  = sum(top5_lastyear, na.rm = T)) %>%
  ungroup() %>%
  summarise(mean25 = mean(carryover25)/25,
            mean10 = mean(carryover10)/10,
            mean5 = mean(carryover5)/5)
```


```{r}
yoy_top <- 
wr %>%
  filter(is.na(league_rank) == FALSE) %>%
  mutate(top25 = ifelse(league_rank <= 25,1,0),
         top10 = ifelse(league_rank <= 10,1,0),
         top5  = ifelse(league_rank <= 5,1,0)) %>%
  arrange(player,year) %>%
  group_by(player) %>%
  mutate(top25_lastyear = ifelse(lag(year) == year - 1, lag(top25),NA),
         top10_lastyear = ifelse(lag(year) == year - 1, lag(top10),NA),
         top5_lastyear  = ifelse(lag(year) == year - 1, lag(top5),NA),
         team_rank_lastyear = ifelse(lag(year) == year - 1,lag(team_rank),NA),
         league_rank_lastyear = ifelse(lag(year) == year - 1,lag(league_rank),NA))

yoy_top %>%
  filter(top25 == 1) %>%
  group_by(year) %>%
  summarise(total_players = sum(top25),
            carryover = sum(top25_lastyear, na.rm = T))
```



## Benchmarks

top 40 YOY will make it the next year


```{r}
accuracy = purrr::map_dbl(.data$models, ~ pROC::auc(pROC::roc(response = insight::get_response(.x), predictor = stats::predict.glm(.x, stats::model.frame(.x)))))


yoy_top %>%
  ungroup() %>%
  filter(is.na(top25_lastyear) == F) %>%
  split(.$year) %>% # from base R
  map_dbl(~pROC::auc(predictor = .$top25_lastyear, response = .$top25))
```

## Build some simple models

```{r}
colnames(yoy_top)
```

```{r}
yoy_top %>%
  ungroup() %>%
  filter(is.na(top25_lastyear) == F) %>%
  split(.$year) %>% # from base R
  map_dbl(~pROC::auc(predictor = glm(data = .,
                                     top25 ~ 
                                       experience + 
                                       age + 
                                       league_rank_lastyear  + 
                                       top25_lastyear +
                                       top10_lastyear + 
                                       top5_lastyear,
                                     family = 'binomial')$fitted.values,
                     response = .$top25))
```

Who did it predict for 2018?

```{r}
m <- yoy_top %>%
  ungroup() %>%
  filter(year == 2018) %>%
  glm(data = .,
      top25 ~ 
        experience + 
        age + 
        league_rank_lastyear  + 
        top25_lastyear +
        top10_lastyear + 
        top5_lastyear,
        family = 'binomial')


df <- 
  yoy_top %>%
  ungroup() %>%
  filter(year == 2018,
         is.na(top25_lastyear) == F)

glimpse(df)

tibble(player = df$name,
       pred = m$fitted.values,
       top25 = df$league_rank) %>%
  arrange(desc(pred))

```

Try a shrinkage model

```{r}
library(glmnet)
features <- c('experience','age','league_rank_lastyear','top40_lastyear','top20_lastyear','top10_lastyear','top5_lastyear')

yoy_top_noNA <- yoy_top[is.na(yoy_top$top40_lastyear) == F,]

lasso <- glmnet(x = as.matrix(yoy_top_noNA[,features]),
                y = yoy_top_noNA$top40,
                family = 'binomial',
                alpha = 1)

which.max(lasso$dev.ratio)
lambda <- lasso$lambda[67]


final_lasso <- glmnet(x = as.matrix(yoy_top_noNA[,features]),
                      y = yoy_top_noNA$top40,
                      family = 'binomial',
                      lambda = lambda,
                      alpha = 1)
?predict.glmnet
predict(final_lasso,newx = as.matrix(yoy_top_noNA[,features]),s =lambda,type='response')

pROC::auc(response = yoy_top_noNA$top40,
          predictor = as.numeric(predict(final_lasso,
                              newx = as.matrix(yoy_top_noNA[,features]),
                              s =lambda,
                              type='response')))


```

```{r}
final_lasso <- glmnet(x = as.matrix(yoy_top_noNA[,features]),
                      y = yoy_top_noNA$top40,
                      family = 'binomial',
                      lambda = lambda,
                      alpha = 1)

final_lasso$lambda[which.max(final_lasso$dev.ratio)]

final_lasso$lambda.1se

```



```{r}
yoy_top %>%
  filter(is.na(top40_lastyear) == F) %>%
  split(.$year) %>%
  map(.,
      ~ glmnet(x = model.matrix(data = .,
                                .$top40 ~ 
                                  .$experience + 
                                  .$age + 
                                  .$league_rank_lastyear  + 
                                  .$top40_lastyear + 
                                  .$top20_lastyear +
                                  .$top10_lastyear + 
                                  .$top5_lastyear)[,-1],
               y = .$top40,
               family = 'binomial',
               alpha = 1,)) %>%
  map(.,
      ~.$lambda[which.max(.$dev.ratio)])

## Ok so this works...now I need to use it. Let's see if we can find the best model 

```


```{r}
# write a function to fit a glmnet, find the best value of alpha and lambda then fit again with those values

yoy_top %>%
  filter(is.na(top40_lastyear) == F) %>%
  split(.$year) %>%
  map(., ~pROC::auc(response = .$top40,
                    predictor = as.numeric(
                      predict(
                        glmnet(x = model.matrix(data = .,
                                                .$top40 ~
                                                  .$experience + 
                                                  .$age + 
                                                  .$league_rank_lastyear  + 
                                                  .$top40_lastyear + 
                                                  .$top20_lastyear +
                                                  .$top10_lastyear + 
                                                  .$top5_lastyear)[,-1],
                               y = .$top40,
                               family = 'binomial',
                               alpha = 1),
                       newx = as.matrix(.[,features]),
                      s=cv.glmnet(x = model.matrix(data = .,
                                                .$top40 ~
                                                  .$experience + 
                                                  .$age + 
                                                  .$league_rank_lastyear  + 
                                                  .$top40_lastyear + 
                                                  .$top20_lastyear +
                                                  .$top10_lastyear + 
                                                  .$top5_lastyear)[,-1],
                               y = .$top40,
                               family = 'binomial',
                               alpha = 1)$lambda.1se,
                      type='response'
                      )
                      )
                    )
      )
```

## Now compare glm to glmnet
```{r}
yoy_top %>%
  ungroup() %>%
  filter(is.na(top40_lastyear) == F) %>%
  split(.$year) %>% # from base R
  map_dbl(~pROC::auc(predictor = glm(data = .,
                                     top40 ~ 
                                       experience + 
                                       age + 
                                       league_rank_lastyear  + 
                                       top40_lastyear + 
                                       top20_lastyear +
                                       top10_lastyear + 
                                       top5_lastyear,
                                     family = 'binomial')$fitted.values,
                     response = .$top40)) -> glm_out

yoy_top %>%
  filter(is.na(top40_lastyear) == F) %>%
  split(.$year) %>%
  map_dbl(., ~pROC::auc(response = .$top40,
                    predictor = as.numeric(
                      predict(
                        glmnet(x = model.matrix(data = .,
                                                .$top40 ~
                                                  .$experience + 
                                                  .$age + 
                                                  .$league_rank_lastyear  + 
                                                  .$top40_lastyear + 
                                                  .$top20_lastyear +
                                                  .$top10_lastyear + 
                                                  .$top5_lastyear)[,-1],
                               y = .$top40,
                               family = 'binomial',
                               alpha = 1),
                       newx = as.matrix(.[,features]),
                      s=cv.glmnet(x = model.matrix(data = .,
                                                .$top40 ~
                                                  .$experience + 
                                                  .$age + 
                                                  .$league_rank_lastyear  + 
                                                  .$top40_lastyear + 
                                                  .$top20_lastyear +
                                                  .$top10_lastyear + 
                                                  .$top5_lastyear)[,-1],
                               y = .$top40,
                               family = 'binomial',
                               alpha = 1)$lambda.1se,
                      type='response'
                      )
                      )
                    )
      ) -> glmnet_out
```

```{r}
tibble(glm_out, glmnet_out) %>%
  mutate(glmnet_pct_better = glmnet_out/glm_out)
```

According to this, glmnet is only better once! Why would glm do so poorly?

For one, it's not predicting on "unseen" data. I think we need to do that. 

Can we use purrr to do that even?

Steps to build models iteratively...

1. glm
2. glmnet
3. xgboost

We also need to incorporate more featuers...
Want to incorporate:
1. QB features
2. First 8 / Last 8 games topX

Want to know...
1. How deep until you capture all you need to capture -- this becomes the input set for predicting who will be in each tier
of the top

Need to fix...the database!



