---
title: "WR Models"
output: html_notebook
---


```{r}
library(tidyverse)
library(xgboost)
load("/home/john/fantasy-football/data/analysis-data/wr.Rda")
load("/home/john/fantasy-football/data/analysis-data/wr_offense.Rda")
wr_model_data <- 
wr %>%
  filter(is.na(league_rank) == FALSE) %>%
  mutate(top40 = ifelse(league_rank <= 40,1,0),
         top20 = ifelse(league_rank <= 20,1,0),
         top10 = ifelse(league_rank <= 10,1,0),
         top5  = ifelse(league_rank <= 5,1,0)) %>%
  arrange(player,year) %>%
  group_by(player) %>%
  mutate(top40_lastyear = ifelse(lag(year) == year - 1, lag(top40),NA),
         top20_lastyear = ifelse(lag(year) == year - 1, lag(top20),NA),
         top10_lastyear = ifelse(lag(year) == year - 1, lag(top10),NA),
         top5_lastyear  = ifelse(lag(year) == year - 1, lag(top5),NA),
         team_rank_lastyear = ifelse(lag(year) == year - 1,lag(team_rank),NA),
         league_rank_lastyear = ifelse(lag(year) == year - 1,lag(league_rank),NA))
```

```{r}

```


#### Models

- We have four outcomes of interest, top 5, 10, 20 and 40. 
- Want to start to predict one year's top X based on previous data.
- use xgboost so we can use NA values

model1: use year 2000 values to predict 2001 values
model2: use year 2000 and 2001 values to predict 2002 values
```{r}
colnames(wr_model_data)
```

```{r}
model1 <- glm(data = wr_model_data %>% filter(year == 2001),
             family = "binomial",
             formula = top20 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear)
summary(model1)

?predict.glm
pred_values1 <- predict.glm(model1, 
                        newdata = wr_model_data %>% filter(year == 2002),
                        type = "response")

df <- data.frame(pv = pred_values1,
                 act = wr_model_data %>% ungroup() %>% filter(year == 2002) %>% select(top20)) %>%
  filter(is.na(pv) == FALSE)

glimpse(df)
str(df)
table(df$act)
pROC::auc(df$top20, df$pv) 
df %>% View()
```



```{r}
pred_auc_values20 <- numeric()
baseline_auc_values20 <- numeric()

for(i in c(2001:2016)){
 

  model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= i),
               family = "binomial",
               formula = top20 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
 df <- data.frame(pv = predict.glm(model, 
                        newdata = wr_model_data %>% filter(year == (i + 1)),
                        type = "response"),
                  act = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top20),
                  last_year = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top20_lastyear)) %>%
     filter(is.na(pv) == FALSE)
                  
  
 pred_auc_values20 <- c(pred_auc_values20, pROC::auc(df$top20, df$pv))
 baseline_auc_values20 <- c(baseline_auc_values20, pROC::auc(df$top20, df$top20_lastyear))
 
 rm(model, df)
 
}

```

```{r}
plot(pred_auc_values20, ylim = c(0.5,1.0))
points(baseline_auc_values20, col = 'red')
```



```{r}
pred_auc_values40 <- numeric()
baseline_auc_values40 <- numeric()

for(i in c(2001:2016)){
 

  model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= i),
               family = "binomial",
               formula = top40 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
 df <- data.frame(pv = predict.glm(model, 
                        newdata = wr_model_data %>% filter(year == (i + 1)),
                        type = "response"),
                  act = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top40),
                  last_year = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top40_lastyear)) %>%
     filter(is.na(pv) == FALSE)
                  
  
 pred_auc_values40 <- c(pred_auc_values40, pROC::auc(df$top40, df$pv))
 baseline_auc_values40 <- c(baseline_auc_values40, pROC::auc(df$top40, df$top40_lastyear))
 
 rm(model, df)
 
}

plot(pred_auc_values40, ylim = c(0.5,1.0))
points(baseline_auc_values40, col = 'red')
```


```{r}
pred_auc_values10 <- numeric()
baseline_auc_values10 <- numeric()

for(i in c(2001:2016)){
 

  model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= i),
               family = "binomial",
               formula = top10 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
 df <- data.frame(pv = predict.glm(model, 
                        newdata = wr_model_data %>% filter(year == (i + 1)),
                        type = "response"),
                  act = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top10),
                  last_year = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top10_lastyear)) %>%
     filter(is.na(pv) == FALSE)
                  
  
 pred_auc_values10 <- c(pred_auc_values10, pROC::auc(df$top10, df$pv))
 baseline_auc_values10 <- c(baseline_auc_values10, pROC::auc(df$top10, df$top10_lastyear))
 
 rm(model, df)
 
}

plot(pred_auc_values10, ylim = c(0.5,1.0))
points(baseline_auc_values10, col = 'red')
```


```{r}
pred_auc_values5 <- numeric()
baseline_auc_values5 <- numeric()

for(i in c(2001:2016)){
 

  model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= i),
               family = "binomial",
               formula = top5 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
 df <- data.frame(pv = predict.glm(model, 
                        newdata = wr_model_data %>% filter(year == (i + 1)),
                        type = "response"),
                  act = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top5),
                  last_year = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top5_lastyear)) %>%
     filter(is.na(pv) == FALSE)
                  
  
 pred_auc_values5 <- c(pred_auc_values5, pROC::auc(df$top5, df$pv))
 baseline_auc_values5 <- c(baseline_auc_values5, pROC::auc(df$top5, df$top5_lastyear))
 
 rm(model, df)
 
}

plot(pred_auc_values5, ylim = c(0.5,1.0))
points(baseline_auc_values5, col = 'red')
```

## Predict 2018 scores...

```{r}
data2018 <- wr_model_data %>%
  filter(year == 2017) %>%
  mutate(experience = experience + 1,
         age = age + 1) %>%
  select(-top40_lastyear,-top20_lastyear,-top10_lastyear,-top5_lastyear,-team_rank_lastyear,-league_rank_lastyear) %>%
  rename(top40_lastyear = top40,
         top20_lastyear = top20,
         top10_lastyear = top10,
         top5_lastyear = top5,
         team_rank_lastyear = team_rank,
         league_rank_lastyear = league_rank)
  
model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= 2017),
               family = "binomial",
               formula = top20 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
summary(model)
 
df <- data.frame(pv = predict.glm(model, 
                        newdata = data2018,
                        type = "response"),
                  data2018$name,
                  data2018$league_rank_lastyear,
                  data2018$team_rank_lastyear)
                  
```


```{r}
data2018 <- wr_model_data %>%
  filter(year == 2017) %>%
  mutate(experience = experience + 1,
         age = age + 1) %>%
  select(-top40_lastyear,-top20_lastyear,-top10_lastyear,-top5_lastyear,-team_rank_lastyear,-league_rank_lastyear) %>%
  rename(top40_lastyear = top40,
         top20_lastyear = top20,
         top10_lastyear = top10,
         top5_lastyear = top5,
         team_rank_lastyear = team_rank,
         league_rank_lastyear = league_rank)
  
model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= 2017),
               family = "binomial",
               formula = top10 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
summary(model)
 
df <- data.frame(pv = predict.glm(model, 
                        newdata = data2018,
                        type = "response"),
                  data2018$name,
                  data2018$league_rank_lastyear,
                  data2018$team_rank_lastyear)

View(df)
```


```{r}
data2018 <- wr_model_data %>%
  filter(year == 2017) %>%
  mutate(experience = experience + 1,
         age = age + 1) %>%
  select(-top40_lastyear,-top20_lastyear,-top10_lastyear,-top5_lastyear,-team_rank_lastyear,-league_rank_lastyear) %>%
  rename(top40_lastyear = top40,
         top20_lastyear = top20,
         top10_lastyear = top10,
         top5_lastyear = top5,
         team_rank_lastyear = team_rank,
         league_rank_lastyear = league_rank)
  
model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= 2017),
               family = "binomial",
               formula = top5 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
summary(model)
 
df <- data.frame(pv = predict.glm(model, 
                        newdata = data2018,
                        type = "response"),
                  data2018$name,
                  data2018$league_rank_lastyear,
                  data2018$team_rank_lastyear)

View(df)
```


```{r}
data2018 <- wr_model_data %>%
  filter(year == 2017) %>%
  mutate(experience = experience + 1,
         age = age + 1) %>%
  select(-top40_lastyear,-top20_lastyear,-top10_lastyear,-top5_lastyear,-team_rank_lastyear,-league_rank_lastyear) %>%
  rename(top40_lastyear = top40,
         top20_lastyear = top20,
         top10_lastyear = top10,
         top5_lastyear = top5,
         team_rank_lastyear = team_rank,
         league_rank_lastyear = league_rank)
  
model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= 2017),
               family = "binomial",
               formula = top40 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
summary(model)
 
df <- data.frame(pv = predict.glm(model, 
                        newdata = data2018,
                        type = "response"),
                  data2018$name,
                  data2018$league_rank_lastyear,
                  data2018$team_rank_lastyear)

View(df)
```



## XGBoost  

```{r}
?xgboost::xgboost

pred_auc_values <- numeric()
baseline_auc_values <- numeric()

for(i in c(2001:2016)){
 

  model <-  glm(data = wr_model_data %>% filter(year >= 2001, year <= i),
               family = "binomial",
               formula = top5 ~ experience + age + top40_lastyear + top20_lastyear + top10_lastyear + top5_lastyear + team_rank_lastyear +
               league_rank_lastyear
               )
 
 df <- data.frame(pv = predict.glm(model, 
                        newdata = wr_model_data %>% filter(year == (i + 1)),
                        type = "response"),
                  act = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top5),
                  last_year = wr_model_data %>% ungroup() %>% filter(year == (i + 1)) %>% select(top5_lastyear)) %>%
     filter(is.na(pv) == FALSE)
                  
  
 pred_auc_values <- c(pred_auc_values, pROC::auc(df$top5, df$pv))
 baseline_auc_values <- c(baseline_auc_values, pROC::auc(df$top5, df$top5_lastyear))
 
 rm(model, df)
 
}
```


```{r}
labels <- c("top5","top10","top20","top40")
etas = c(0.01,0.1,0.2,0.3)
depths = c(1,2,4)
trees = c(32,64,128)

feature_names <- c("experience","age","top40_lastyear","top20_lastyear","top10_lastyear","top5_lastyear",
                                         "team_rank_lastyear","league_rank_lastyear")

j <- 3
l <- 3
k <- 1
m <- 1
model_data <- wr_model_data %>% 
      ungroup() %>%
      filter(year >= 2001, year <= 2010) %>%
      select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()

    ix <- sample(c(0,1), dim(model_data)[1], replace = TRUE, prob = c(1/3,2/3)) 

    
    dtrain <- xgb.DMatrix(model_data[ix==1,c(feature_names)],
                          label=model_data[ix == 1, labels[j]])
    dtest <- xgb.DMatrix(model_data[ix==0,c(feature_names)],
                         label=model_data[ix == 0,labels[j]])
    watchlist <- list(train=dtrain, test=dtest)
    model <- xgb.train(data = dtrain,
                          max.depth = 4, 
                          eta = 0.1,
                          nrounds = 1000, 
                          watchlist = watchlist,
                          eval.metric = "logloss",
#                          eval.metric = "auc",
                          early_stopping_rounds = 10,
                          gamma = 0.00001,
                          objective = "binary:logistic",
#                          maximize = TRUE,
                          verbose = 1)
str(model)
str[2]
?xgboost
model
model$best_iteration

d2011 <- wr_model_data %>%
  ungroup() %>%
  filter(year == 2011) %>%
        select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()
colnames(d2011)  
preds <- predict(model, d2011[,c(feature_names)])
labels[j]
head(d2011)
pdf <- cbind(preds,d2011[,4],wr_model_data %>% ungroup %>% filter(year == 2011) %>% select(name,league_rank))
preds
xgb.importance(model=model)
class(model)

pROC::auc(pdf$`d2011[, 4]`,pdf$preds)
model$best_ntreelimit
preds
```




```{r}
i <- 2010
j <- 1
k <- 1
l <- 1

    model_data <- wr_model_data %>% 
      ungroup() %>%
      filter(year >= 2001, year <= i) %>%
      select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()

    pred_data <- wr_model_data %>%
                  ungroup() %>%
                  filter(year == (i+1)) %>%
                  select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
                  as.matrix()

    ix <- sample(c(0,1), dim(model_data)[1], replace = TRUE, prob = c(1/3,2/3)) 

    dtrain <- xgb.DMatrix(model_data[ix==1,c(feature_names)],
                          label=model_data[ix == 1, labels[j]])
    dtest <- xgb.DMatrix(model_data[ix==0,c(feature_names)],
                         label=model_data[ix == 0,labels[j]])
    watchlist <- list(train=dtrain, test=dtest)
    
    
    
    model <- xgb.train(data = dtrain,
                            max.depth = depths[l], 
                            eta = etas[k],
                            nrounds = 1000, 
                            watchlist = watchlist,
                            eval.metric = "logloss",
                          early_stopping_rounds = 5,
                          gamma = 0.00001,
                          objective = "binary:logistic",
                          verbose = 0)
          
          preds <- predict(model, pred_data[,c(feature_names)])
          pdf <- cbind(pred_data[,j+1],preds,wr_model_data %>% ungroup %>% filter(year == i+1) %>% select(name,league_rank))
          colnames(pdf) <- c("label","preds","name","league_rank")
          results <- bind_rows(results,c(label = labels[j],
                                         eta = etas[k],
                                         depths = depths[l],
                                         auc = pROC::auc(pdf$label,pdf$preds)))    
          counter <- counter + 1
          print(counter)
          rm(model,preds,pdf)  

          
          
preds
pdf

str(
c(label = labels[j],
                                         eta = etas[k],
                                         depths = depths[l],
                                         auc = pROC::auc(pdf$label,pdf$preds))
)
str(results)
```


```{r}
set.seed(2018)
labels <- c("top5","top10","top20","top40")
etas <- c(0.01,0.1,0.2,0.3)
depths <- c(1,2,4,8)

feature_names <- c("experience","age","top40_lastyear","top20_lastyear","top10_lastyear","top5_lastyear",
                                         "team_rank_lastyear","league_rank_lastyear")


results <- data.frame(label = as.character(NULL),
                      eta = NULL,
                      depth = NULL,
                      auc = NULL,
                      year = NULL)
counter <- 1
for(i in c(2001:2016)){
  for(j in seq_along(labels)){
    
    model_data <- wr_model_data %>% 
      ungroup() %>%
      filter(year >= 2001, year <= i) %>%
      select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()

    pred_data <- wr_model_data %>%
                  ungroup() %>%
                  filter(year == (i+1)) %>%
                  select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
                  as.matrix()

    ix <- sample(c(0,1), dim(model_data)[1], replace = TRUE, prob = c(1/3,2/3)) 

    dtrain <- xgb.DMatrix(model_data[ix==1,c(feature_names)],
                          label=model_data[ix == 1, labels[j]])
    dtest <- xgb.DMatrix(model_data[ix==0,c(feature_names)],
                         label=model_data[ix == 0,labels[j]])
    watchlist <- list(train=dtrain, test=dtest)

    for(k in seq_along(etas)){
      for(l in seq_along(depths)){
          model <- xgb.train(data = dtrain,
                            max.depth = depths[l], 
                            eta = etas[k],
                            nrounds = 1000, 
                            watchlist = watchlist,
                            eval.metric = "logloss",
                          early_stopping_rounds = 5,
                          gamma = 0.00001,
                          objective = "binary:logistic",
                          verbose = 0)
          
          preds <- predict(model, pred_data[,c(feature_names)])
          pdf <- cbind(pred_data[,j+1],preds,wr_model_data %>% ungroup %>% filter(year == i+1) %>% select(name,league_rank))
          colnames(pdf) <- c("label","preds","name","league_rank")
          results <- bind_rows(results,c(label = labels[j],
                                         eta = etas[k],
                                         depths = depths[l],
                                         auc = pROC::auc(pdf$label,pdf$preds),
                                         year = i))    
          counter <- counter + 1
          rm(model,preds,pdf)  
      }
    }
    
  }
}

```
```{r}
results
```

```{r}
p <- ggplot(data = results,
            aes(x = label, y = as.numeric(auc), shape = eta, colour = as.factor(depths)))
p + geom_point()
```

```{r}
results %>%
  mutate(auc = as.numeric(auc),
         eta = as.numeric(eta),
         depths = as.numeric(depths)) %>%
  group_by(label,year) %>%
  arrange(desc(auc)) %>%
  dplyr::slice(1)
```

```{r}
results %>%
  mutate(auc = as.numeric(auc),
         eta = as.numeric(eta),
         depths = as.numeric(depths)) %>%
  group_by(label,year) %>%
  arrange(desc(auc)) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  group_by(label,eta,depths) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```


```{r}
best <- results %>%
  mutate(auc = as.numeric(auc),
         eta = as.numeric(eta),
         depths = as.numeric(depths)) %>%
  group_by(label,year) %>%
  arrange(desc(auc)) %>%
  dplyr::slice(1)

best
```


Try using data that is at most 5 years old. That way any recent trends are eliminated. 


```{r}
set.seed(2018)
labels <- c("top5","top10","top20","top40")
etas <- c(0.01,0.1,0.2,0.3)
depths <- c(1,2,4,8)

feature_names <- c("experience","age","top40_lastyear","top20_lastyear","top10_lastyear","top5_lastyear",
                                         "team_rank_lastyear","league_rank_lastyear")


last_five_results <- data.frame(label = as.character(NULL),
                      eta = NULL,
                      depth = NULL,
                      auc = NULL,
                      year = NULL)
counter <- 1
for(i in c(2001:2016)){
  for(j in seq_along(labels)){
    
    model_data <- wr_model_data %>% 
      ungroup() %>%
      filter(year >= (i - 4), year <= i) %>%
      select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()

    pred_data <- wr_model_data %>%
                  ungroup() %>%
                  filter(year == (i+1)) %>%
                  select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
                  as.matrix()

    ix <- sample(c(0,1), dim(model_data)[1], replace = TRUE, prob = c(1/3,2/3)) 

    dtrain <- xgb.DMatrix(model_data[ix==1,c(feature_names)],
                          label=model_data[ix == 1, labels[j]])
    dtest <- xgb.DMatrix(model_data[ix==0,c(feature_names)],
                         label=model_data[ix == 0,labels[j]])
    watchlist <- list(train=dtrain, test=dtest)

    for(k in seq_along(etas)){
      for(l in seq_along(depths)){
          model <- xgb.train(data = dtrain,
                            max.depth = depths[l], 
                            eta = etas[k],
                            nrounds = 1000, 
                            watchlist = watchlist,
                            eval.metric = "logloss",
                          early_stopping_rounds = 5,
                          gamma = 0.00001,
                          objective = "binary:logistic",
                          verbose = 0)
          
          preds <- predict(model, pred_data[,c(feature_names)])
          pdf <- cbind(pred_data[,j+1],preds,wr_model_data %>% ungroup %>% filter(year == i+1) %>% select(name,league_rank))
          colnames(pdf) <- c("label","preds","name","league_rank")
          results <- bind_rows(results,c(label = labels[j],
                                         eta = etas[k],
                                         depths = depths[l],
                                         auc = pROC::auc(pdf$label,pdf$preds),
                                         year = i))    
          counter <- counter + 1
          rm(model,preds,pdf)  
      }
    }
    
  }
}
```


```{r}
best_last5 <- results %>%
  group_by(label,year) %>%
  mutate(auc = as.numeric(auc)) %>%
  arrange(desc(auc)) %>%
  dplyr::slice(1)
best_last5
```

```{r}
all_model_results <-
  best_last5 %>%
  ungroup() %>%
  mutate(year = as.numeric(year),
         model = "xgb_last5") %>%
  select(label,year,auc,model) %>%
  bind_rows(best %>%
              ungroup() %>%
              mutate(year = as.numeric(year),
                     model = "xgb") %>%
              select(label,year,auc,model)
  ) %>%
  bind_rows(
    tibble(label = rep("top5",16),
           year = c(2001:2016),
           auc = pred_auc_values5,
           model = "glm")
  ) %>%
    bind_rows(
    tibble(label = rep("top5",16),
           year = c(2001:2016),
           auc = baseline_auc_values5,
           model = "baseline")
  ) %>%
    bind_rows(
    tibble(label = rep("top10",16),
           year = c(2001:2016),
           auc = pred_auc_values10,
           model = "glm")
  ) %>%
    bind_rows(
    tibble(label = rep("top10",16),
           year = c(2001:2016),
           auc = baseline_auc_values10,
           model = "baseline")
  ) %>%
    bind_rows(
    tibble(label = rep("top20",16),
           year = c(2001:2016),
           auc = pred_auc_values20,
           model = "glm")
  ) %>%
    bind_rows(
    tibble(label = rep("top20",16),
           year = c(2001:2016),
           auc = baseline_auc_values20,
           model = "baseline")
  ) %>%
    bind_rows(
    tibble(label = rep("top40",16),
           year = c(2001:2016),
           auc = pred_auc_values40,
           model = "glm")
  ) %>%
    bind_rows(
    tibble(label = rep("top40",16),
           year = c(2001:2016),
           auc = baseline_auc_values40,
           model = "baseline")
  ) 
  
  
```

```{r}
all_model_results %>%
  ggplot(data = .,
         aes(x = year,
             y = auc,
             group = model,
             colour = model)) +
  geom_point() +
  geom_line() +
  geom_smooth(se = F) +
  facet_wrap(~label,ncol = 1)
```

Build a model using all data from 2017 and before...
Use validation data to determine hyperparameters...
Predict 2018 figures

```{r}
set.seed(2018)
labels <- c("top5","top10","top20","top40")
etas <- c(0.01,0.1,0.2,0.3)
depths <- c(1,2,4,8)

feature_names <- c("experience","age","top40_lastyear","top20_lastyear","top10_lastyear","top5_lastyear",
                                         "team_rank_lastyear","league_rank_lastyear")


models <- list()

counter <- 1
i <- 2017
  for(j in seq_along(labels)){
    models[[labels[j]]] <- list()
    model_counter <- 1
    model_data <- wr_model_data %>% 
      ungroup() %>%
#      filter(year >= (i - 4), year <= i) %>%
      select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
      as.matrix()

    ix <- sample(c(0,1), dim(model_data)[1], replace = TRUE, prob = c(1/3,2/3)) 

    dtrain <- xgb.DMatrix(model_data[ix==1,c(feature_names)],
                          label=model_data[ix == 1, labels[j]])
    dtest <- xgb.DMatrix(model_data[ix==0,c(feature_names)],
                         label=model_data[ix == 0,labels[j]])
    watchlist <- list(train=dtrain, test=dtest)

    for(k in seq_along(etas)){
      for(l in seq_along(depths)){
          models[[labels[[j]]]][[model_counter]] <- 
            xgb.train(data = dtrain,
                      max.depth = depths[l], 
                      eta = etas[k],
                      nrounds = 1000, 
                      watchlist = watchlist,
                      eval.metric = "logloss",
                      early_stopping_rounds = 5,
                      gamma = 0.00001,
                      objective = "binary:logistic",
                      verbose = 0)
     
          model_counter <- model_counter + 1

      }
    }
    
  }



```

```{r}




models[[1]][[1]]$params$max_depth
models[[1]][[1]]$params$eta
models[[1]][[1]]$params$gamma

models[[1]][[1]]$best_score

## Extract the values of max_depth, eta, and gamma where best_score is lowest for each of the 16 models in each element.

which.min(models[[1]][[1:16]]$best_score)

for(i in c(1:4)){
  best_ix <- which.min(lapply(models[[i]],"[[","best_score"))
  print(best_ix)
print(c(labels[[i]],
  models[[i]][[best_ix]]$params$max_depth,
  models[[i]][[best_ix]]$params$eta,
  models[[i]][[best_ix]]$params$gamma))
}





```

Build 2018 prediction data
```{r}
name <- wr_model_data %>%
        ungroup() %>%
        filter(year == 2017) %>%
        select(name) %>%
  unlist()
name
  
pred_data <-
wr_model_data %>%
  filter(year == 2017) %>%
  ungroup() %>%
  mutate(age = age + 1,
         experience = experience + 1,
         top40_lastyear = top40,
         top20_lastyear = top20,
         top10_lastyear = top10,
         top5_lastyear = top5,
         team_rank_lastyear = team_rank,
         league_rank_lastyear = league_rank) %>%
  select(year,top5,top10,top20,top40,experience,age,top40_lastyear,top20_lastyear,top10_lastyear,top5_lastyear,
                                         team_rank_lastyear,league_rank_lastyear) %>%
                  as.matrix()
pred_data %>% as.data.frame() #%>% select(name, league_rank_lastyear)

head(pred_data)

preds2018 <- 
  tibble(name,
      top5 = predict(models[[1]][[10]], pred_data[,c(feature_names)]),
      top10 = predict(models[[2]][[10]], pred_data[,c(feature_names)]),
      top20 = predict(models[[3]][[14]], pred_data[,c(feature_names)]),
      top40 = predict(models[[4]][[14]], pred_data[,c(feature_names)])) %>%
  left_join(tibble(name = name, pred_data[,13]),
            by = "name")



```


```{r}
head(preds2018)
cor(preds2018[,c(2:4)])


ggplot(data = as.tibble(preds2018),
       aes(x = top5, y = top40,label=name)) +
  geom_point() +
  geom_label()

```

