---
title: "Clustering Cost and Production by ADP"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(DBI)
library(pool)
```


There are three data sources to do this:
1. Cost (Draft data)
2. Production (From Armchair Analysis DB)
3. ADP data (scraped from the web)

Draft data
```{r}
load(file = "/home/john/projects/fantasy-football/data/bn-draft-data/draft_data.Rda")
```

Production Data
```{r}
pool <- dbPool(drv = RMySQL::MySQL(), 
               dbname = "armchair_analysis", 
               host = "localhost", 
               username = "john", 
               port = 3306, 
               unix.sock = "/var/run/mysqld/mysqld.sock")
team_name_abbr_lookup <- data.frame(abbr = c('ARI','ATL','BAL','BUF','CAR','CHI','CIN','CLE','DAL','DEN',
                                             'DET','GB' ,'HOU','IND','JAC','KC' ,'LA' ,'MIA','MIN','NE',
                                             'NO' ,'NYG','NYJ','OAK','PHI','PIT','SD' ,'SEA','SF' , 'STL',
                                             'TB','TEN','WAS'),
                                    name = c('Cardinals','Falcons','Ravens','Bills','Panthers','Bears','Bengals','Browns',
                                             'Cowboys','Broncos','Lions','Packers','Texans','Colts','Jaguars','Chiefs','Rams',
                                             'Dolphins','Vikings','Patriots','Saints','Giants','Jets','Raiders','Eagles',
                                             'Steelers','Chargers','Seahawks','49ers','Rams','Buccaneers','Titans','Redskins'),
                                    stringsAsFactors = FALSE)

offense <- dbGetQuery(con = pool,
                      "SELECT 
                      a.player
                      , b.wk as week
                      , a.year
                      , a.team
                      , 0.04*py + -2*ints + 4*tdp + 0.1*ry + 0.1*recy + 6*tdr + 6*tdrec -2*fuml +6*tdret as fp
                      ,py
                      ,ints
                      ,tdp
                      ,ry
                      ,recy
                      ,tdr
                      ,tdrec
                      ,fuml
                      ,tdret
                      FROM offense as a
                      JOIN game as b on a.gid = b.gid") 


two_point <- dbGetQuery(con = pool,
                        "select 
                        a.*
                        , b.gid
                        , c.seas as year
                        , c.wk as week 
                        from conv as a 
                        join play as b on a.pid = b.pid
                        join game as c on b.gid = c.gid where conv = 1")


rush_two_point <- two_point %>%
  filter(bc != "")
pass_two_point <- two_point %>%
  filter(psr != "")
two_point %>%
  filter(bc == "", psr == "") # forget this one


offense2 <-
  offense %>%
  left_join(., 
            pass_two_point %>% group_by(week,year,psr) %>% summarise(count = n()), 
            by = c("year","week","player" = "psr")) %>% 
  mutate(fp = ifelse(is.na(count) == FALSE, fp + 2*count, fp)) %>%
  select(-count)

offense3 <-
  offense2 %>%
  left_join(.,
            pass_two_point %>% group_by(week,year,trg) %>% summarise(count= n()),
            by = c("year","week","player" = "trg")) %>%
  mutate(fp = ifelse(is.na(count) == FALSE, fp+2*count, fp)) %>%
  select(-count)

offense4 <-
  offense3 %>%
  left_join(.,
            rush_two_point %>% group_by(week,year,bc) %>% summarise(count = n()),
            by = c("year","week","player" = "bc")) %>%
  mutate(fp = ifelse(is.na(count) == FALSE, fp+2*count, fp)) %>%
  select(-count)

offense5 <-
  offense4 %>%
  filter(week <= 16) %>%
  group_by(player,year) %>%
  summarise(ppg= mean(fp),
            ppg_sd = sd(fp),
            gp = n()) %>%
  filter(gp >= 4)


positions <- dbGetQuery(con = pool,
                  "SELECT 
                   a.player
                  ,a.fname
                  ,a.lname
                  ,a.dob
                  ,a.start
                  ,a.pos1 as position
                  FROM player AS a")

offense6 <- offense5 %>%
  left_join(positions, by = "player")


```



ADP Data
```{r}
load("/home/john/projects/fantasy-football/data/adp-data/adp.Rda")
```



Need a Dataset with
|player|year|pos|pos_adp|cost|fp|


```{r}
glimpse(adp)
```

```{r}
glimpse(draft_data)
```

```{r}
glimpse(offense5)
```

```{r}
cost_adp_prod <-
offense6 %>%
  select(player,year,position,ppg,ppg_sd) %>%
  inner_join(adp %>% select(year,player_code,pos_adp),
             by = c("player" = "player_code","year")) %>%
  left_join(draft_data %>% select(year,player_code,player,keeper,adj_value),
             by = c("player" = "player_code","year")) 

save(cost_adp_prod,file="/home/john/projects/fantasy-football/data/clustering-data/cost_adp_prod.Rda")
```

```{r}
glimpse(draft_data)
```




```{r}
glimpse(offense5)
```


```{r}
cost_adp_prod %>%
  filter(pos == 'QB', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```


```{r}
cost_adp_prod %>%
  filter(pos == 'RB', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```


```{r}
cost_adp_prod %>%
  filter(pos == 'WR', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```

```{r}
cost_adp_prod %>%
  filter(pos == 'TE', keeper == 0) %>%
  ggplot(data = .,
         aes(x = adj_value, y = ppg,label = pos_adp)) +
  geom_label()
```

```{r}
table(cost_adp_prod$pos)
```

center and scale ppg, ppg_sd, and cost
```{r}
head(cost_adp_prod)
```

Remove keepers and Andre Johnson

```{r}
sum(is.na(cost_adp_prod$adj_value))
```



```{r}
cost_adp_prod_cleaned <- 
  cost_adp_prod %>% 
  filter(keeper == 0, adj_value < 200) %>%
  mutate(adj_value = ifelse(is.na(adj_value),0,adj_value))
```

```{r}

```


```{r}
to_scale_vars <- cost_adp_prod_cleaned[,c(3,4,9)]
to_scale_vars_scaled <- as.data.frame(scale(to_scale_vars))
colnames(to_scale_vars_scaled) 
colnames(to_scale_vars_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(to_scale_vars_scaled)
#var(to_scale_vars_centered[,3])
```

```{r}
cost_adp_prod_scaled_cleaned <- bind_cols(cost_adp_prod_cleaned,to_scale_vars_scaled)
save(cost_adp_prod_scaled_cleaned,
     file = "/home/john/projects/fantasy-football/data/clustering-data/cost_adp_prod_scaled_cleaned.Rda")
```

```{r}
head(cost_adp_prod_scaled_cleaned)
```


### Make a version of the data where I impute the averaages cost for adp for all kept values


Stil exclude Andre Johnson...

```{r}
cost_adp_prod_no_keepers <- 
  cost_adp_prod %>% 
  filter((is.na(keeper) | keeper == 0), (is.na(adj_value) | adj_value < 200),year > 2010) %>%
  mutate(adj_value = ifelse(is.na(adj_value),0,adj_value))

sum(is.na(cost_adp_prod$adj_value))
sum(is.na(cost_adp_prod_no_keepers$adj_value))
sum(is.na(cost_adp_prod$keeper))
sum(cost_adp_prod_no_keepers$adj_value == 0)

adp_pos_average_cost <-
  cost_adp_prod_no_keepers %>%
  group_by(position,pos_adp) %>%
  summarise(mean_cost = mean(adj_value),
            num_zero = sum(adj_value == 0))


```




```{r}
cost_adp_prod_keepers <- 
  cost_adp_prod %>%
  filter(keeper == 1, year >= 2010) %>%
  left_join(adp_pos_average_cost, by = c('pos_adp','position')) %>%
  mutate(adj_value = mean_cost)
  
```

```{r}
which(is.na(cost_adp_prod_keepers$adj_value))
```

```{r}
cost_adp_prod_keepers[67,]
```


#6 RBs are not available...take the average of #s 5 and 7...

```{r}
cost_adp_prod_no_keepers %>%
  filter(position == 'QB', pos_adp %in% c(24,26))
```

```{r}
# use a weighted average
(3*(80+2/3) + 4*(63+2/3)) /7
```

```{r}
cost_adp_prod_keepers2 <-
  cost_adp_prod_keepers %>%
  mutate(adj_value = ifelse((position == 'QB' & pos_adp == 25),(18+1)/2,adj_value))
```



```{r}
colnames(cost_adp_prod_no_keepers)
```

```{r}
colnames(cost_adp_prod_keepers)
```


```{r}
imputed_cost_adp_prod <- 
  bind_rows(cost_adp_prod_no_keepers,
            cost_adp_prod_keepers2 %>%
              select(1,2,3,4,5,6,7,8,9))
```

```{r}
colnames(imputed_cost_adp_prod)
```



```{r}
to_scale_vars <- imputed_cost_adp_prod[,c(4,5,9)]
to_scale_vars_scaled <- as.data.frame(scale(to_scale_vars))
head(to_scale_vars_scaled)
lapply(to_scale_vars_scaled, function(x) sum(is.na(x)))
which(is.na(to_scale_vars_scaled[,3]))
#imputed_cost_adp_prod[c(788,834,836,847,850),]
colnames(to_scale_vars_scaled) 
colnames(to_scale_vars_scaled) <- c('ppg_scaled','ppg_sd_scaled','adj_value_scaled')
colMeans(to_scale_vars_scaled)
var(to_scale_vars_scaled[,3])

imputed_cost_adp_prod_scaled <- bind_cols(imputed_cost_adp_prod,to_scale_vars_scaled)
head(imputed_cost_adp_prod_scaled)
save(cost_adp_prod_scaled_cleaned,
     file = "/home/john/projects/fantasy-football/data/clustering-data/imputed_cost_adp_prod_scaled.Rda")
```

## Plots of Production vs. Cost
 
```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'QB'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()
```


```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'RB'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'WR'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'TE'),
       aes(x = pos_adp, y = ppg)) +
  geom_point() +
  geom_smooth()
```


```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'QB'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()
```


```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'RB'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'WR'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()
```


```{r}
ggplot(data = cost_adp_prod_scaled_cleaned %>% filter(pos == 'TE'),
       aes(x = adj_value, y = ppg)) +
  geom_point() +
  geom_smooth()
```



