---
title: "Wide Receiver Analysis""
output: html_notebook
---

Questions to answer...

Looking for really straightforward signals...


1. By year, how many #1s before a #2
2. By year, how many #2s before a #3
3. By year, how many #3s before a #4
4. Where there is draft data available, how do prices look by league rank and team rank?
5. How well can we predict #1 WRs based on last year's rankings?
6. If a former #1 WR is still on a team, how does that depress other WRs, including subsequent #1s?
7. On a team, what is the lowest point total such that #1 receiver from prior year is on team but is no longer the #1 WR


## Load data and libraries

```{r}
library(tidyverse)
load("/home/john/projects/fantasy-football/data/analysis-data/wr.Rda")
load("/home/john/projects/fantasy-football/data/analysis-data/wr_offense.Rda")
```


## Questions 1-3


#### How many #1s before a number 2
```{r}
wr %>%
  group_by(year) %>%
  arrange(league_rank) %>%
  filter(team_rank == 2) %>%
  slice(1) %>%
  select(league_rank) %>%
  plot()
```

```{r}
wr %>%
  group_by(year) %>%
  arrange(league_rank) %>%
  filter(team_rank == 2) %>%
  slice(1) %>%
  select(league_rank) %>%
  ungroup() %>%
  summarise(min = min(league_rank),
            max = max(league_rank),
            median = median(league_rank),
            mean = mean(league_rank))
```

What are the average league ranks of a #1 and a #2?
```{r}
wr %>%
  group_by(year) %>%
  filter(team_rank == 1) %>%
  summarise(min = min(league_rank),
            max = max(league_rank),
            median = median(league_rank),
            mean = mean(league_rank))
```

```{r}
wr %>%
  group_by(year) %>%
  filter(team_rank == 2) %>%
  summarise(min = min(league_rank),
            max = max(league_rank),
            median = median(league_rank),
            mean = mean(league_rank))
```

```{r}
wr %>%
  filter(team_rank %in% c(1,2,3)) %>%
  group_by(year,team_rank) %>%
  summarise(median = median(total_points)/16,
            mean = mean(total_points)/16) %>%
  ggplot(data = .,
         aes(x = year, y = median, group = team_rank, colour = team_rank)) +
  geom_line()
```

```{r}
wr %>%
  filter(team_rank %in% c(1,2,3)) %>%
  group_by(year,team_rank) %>%
  summarise(median = quantile(total_points,0.75)/16,
            mean = quantile(total_points,0.75)/16) %>%
  ggplot(data = .,
         aes(x = year, y = median, group = team_rank, colour = team_rank)) +
  geom_line()
```

```{r}
wr %>%
  filter(team_rank %in% c(1,2,3)) %>%
  group_by(year,team_rank) %>%
  summarise(median = quantile(total_points,0.25)/16,
            mean = quantile(total_points,0.25)/16) %>%
  ggplot(data = .,
         aes(x = year, y = median, group = team_rank, colour = team_rank)) +
  geom_line()
```

## How many #1s and #2s before a #3?

```{r}
wr %>%
  group_by(year) %>%
  arrange(league_rank) %>%
  filter(team_rank == 3) %>%
  slice(1) %>%
  select(league_rank) %>%
  plot()
```

```{r}
wr %>%
  group_by(year) %>%
  arrange(league_rank) %>%
  filter(team_rank == 3) %>%
  slice(1) %>%
  select(league_rank) %>%
  ungroup() %>%
  summarise(min = min(league_rank),
            max = max(league_rank),
            median = median(league_rank),
            mean = mean(league_rank))
```

Who was the 11?

```{r}
wr %>%
  group_by(year) %>%
  arrange(league_rank) %>%
  filter(team_rank == 3) %>%
  slice(1)
```

Look at the distribution by group

```{r}
wr %>%
  filter(team_rank %in% c(1,2,3),
         year >= 2010) %>%
  ggplot(data = .,
         aes(x = team_rank, y = total_points/16, group = team_rank)) +
  geom_boxplot() +
  geom_jitter()
```

## By year, what is the total points cutoff for being a top 20 WR
```{r}
wr %>%
  group_by(year) %>%
  arrange(desc(total_points)) %>%
  slice(20) %>%
  select(name, league_rank, total_points)

```



## Where there is draft data available, how do prices look by league rank and team rank?
```{r}
load("/home/john/fantasy-football/data/bn-draft-data/draft_data.Rda")
glimpse(draft_data)

wr_draft <- full_join(wr,draft_data, by = c('player' = 'player_code','year'))
```

```{r}
ggplot(data = wr_draft %>% filter(keeper == 0, team_rank <= 3, league_rank <= 50),
       aes(x = league_rank, y = adj_value, colour = as.factor(team_rank))) +
  geom_point()
```
```{r}
ggplot(data = wr_draft %>% filter(keeper == 0, team_rank <= 3),
       aes(x = league_rank, y = adj_value, colour = as.factor(team_rank))) +
  geom_point() +
  facet_wrap(~year, ncol = 8) +
  geom_vline(xintercept = 12) +
  geom_text(data = wr_draft %>% filter(keeper == 0, team_rank == 2, league_rank <= 20),
            aes(x = league_rank, y = adj_value, label = name))
```

```{r}
head(wr_draft)
```

```{r}
min(draft_data$year)
which(is.na(wr_draft$adj_value))
```

```{r}
wr_draft %>%
  filter(year == 2010, (is.na(keeper) | keeper == 0)) %>%
  mutate(adj_value = ifelse(is.na(adj_value),-1,adj_value)) %>%
  ggplot(data = .,
         aes(x = adj_value, y = total_points)) +
  geom_point()
```

```{r}
head(wr)
```

```{r}
top20_by_year <- 
wr_draft %>%
  filter(keeper == 0 | is.na(keeper),
         is.na(league_rank) == FALSE) %>%
  mutate(adj_value = ifelse(is.na(adj_value),-1,adj_value)) %>%
  mutate(bin = case_when(
    adj_value == -1 ~ 'undrafted',
    adj_value == 1 ~ '$1',
    adj_value >= 2 & adj_value <= 5 ~ '$2-$5',
    adj_value >= 6 & adj_value <= 15 ~ '$6-$15',
    adj_value >= 16 & adj_value <= 25 ~ '$16-$25',
    adj_value >= 26 & adj_value <= 35 ~ '$26-$35',
    adj_value >= 36 & adj_value <= 50 ~ '$36-$50',
    adj_value >=51 & adj_value <= 75 ~ '$51-$75',
    adj_value > 75 ~ '$75+',
    TRUE ~ as.character(NA)
  )) %>% 
  mutate(top20 = ifelse(league_rank <= 20, 1, 0)) %>%
  group_by(year,bin) %>% 
  summarise(total = n(),
            total_top20 = sum(top20)) %>%
  mutate(pct_top20 = total_top20/total) %>% 
  ungroup() %>%
  filter(year >= 2010) %>%
  arrange(year,bin)
```



```{r}
ggplot(data = top20_by_year %>% mutate(fraction = paste0(total_top20,"/",total)),
       aes(x = bin, y = pct_top20, colour = as.factor(year))) +
  geom_text(aes(x = bin, y = pct_top20, label = fraction))

```

```{r}
top20_by_year %>%
  group_by(bin) %>%
  summarise(mean_pct = mean(pct_top20))
```

```{r}
wr_draft %>%
  filter(keeper == 0 | is.na(keeper),
         is.na(league_rank) == FALSE,
         adj_value < 200) %>% 
    mutate(top20 = ifelse(league_rank <= 20, 1, 0)) %>%
  ggplot(data = .,
         aes(x = adj_value, y = top20)) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = c(26,50))
  
```

Try it for top 10

```{r}
top10_by_year <- 
wr_draft %>%
  filter(keeper == 0 | is.na(keeper),
         is.na(league_rank) == FALSE) %>%
  mutate(adj_value = ifelse(is.na(adj_value),-1,adj_value)) %>%
  mutate(bin = case_when(
    adj_value == -1 ~ 'undrafted',
    adj_value == 1 ~ '$1',
    adj_value >= 2 & adj_value <= 5 ~ '$2-$5',
    adj_value >= 6 & adj_value <= 15 ~ '$6-$15',
    adj_value >= 16 & adj_value <= 25 ~ '$16-$25',
    adj_value >= 26 & adj_value <= 35 ~ '$26-$35',
    adj_value >= 36 & adj_value <= 50 ~ '$36-$50',
    adj_value >=51 & adj_value <= 75 ~ '$51-$75',
    adj_value > 75 ~ '$75+',
    TRUE ~ as.character(NA)
  )) %>% 
  mutate(top10 = ifelse(league_rank <= 10, 1, 0)) %>%
  group_by(year,bin) %>% 
  summarise(total = n(),
            total_top10 = sum(top10)) %>%
  mutate(pct_top10 = total_top10/total) %>% 
  ungroup() %>%
  filter(year >= 2010) 
```

```{r}
ggplot(data = top10_by_year %>% mutate(fraction = paste0(total_top10,"/",total)),
       aes(x = bin, y = pct_top10, colour = as.factor(year))) +
  geom_text(aes(x = bin, y = pct_top10, label = fraction))

```
```{r}
top10_by_year %>%
  group_by(bin) %>%
  summarise(mean_pct = mean(pct_top10))
```

## Do top 5
```{r}
top5_by_year <- 
wr_draft %>%
  filter(keeper == 0 | is.na(keeper),
         is.na(league_rank) == FALSE) %>%
  mutate(adj_value = ifelse(is.na(adj_value),-1,adj_value)) %>%
  mutate(bin = case_when(
    adj_value == -1 ~ 'undrafted',
    adj_value == 1 ~ '$1',
    adj_value >= 2 & adj_value <= 5 ~ '$2-$5',
    adj_value >= 6 & adj_value <= 15 ~ '$6-$15',
    adj_value >= 16 & adj_value <= 25 ~ '$16-$25',
    adj_value >= 26 & adj_value <= 35 ~ '$26-$35',
    adj_value >= 36 & adj_value <= 50 ~ '$36-$50',
    adj_value >=51 & adj_value <= 75 ~ '$51-$75',
    adj_value > 75 ~ '$75+',
    TRUE ~ as.character(NA)
  )) %>% 
  mutate(top5 = ifelse(league_rank <= 5, 1, 0)) %>%
  group_by(year,bin) %>% 
  summarise(total = n(),
            total_top5 = sum(top5)) %>%
  mutate(pct_top5 = total_top5/total) %>% 
  ungroup() %>%
  filter(year >= 2010) 
ggplot(data = top5_by_year %>% mutate(fraction = paste0(total_top5,"/",total)),
       aes(x = bin, y = pct_top5, colour = as.factor(year))) +
  geom_text(aes(x = bin, y = pct_top5, label = fraction))

top5_by_year %>%
  group_by(bin) %>%
  summarise(mean_pct = mean(pct_top5))
```

## top 40
```{r}
top40_by_year <- 
wr_draft %>%
  filter(keeper == 0 | is.na(keeper),
         is.na(league_rank) == FALSE) %>%
  mutate(adj_value = ifelse(is.na(adj_value),-1,adj_value)) %>%
  mutate(bin = case_when(
    adj_value == -1 ~ 'undrafted',
    adj_value == 1 ~ '$1',
    adj_value >= 2 & adj_value <= 5 ~ '$2-$5',
    adj_value >= 6 & adj_value <= 15 ~ '$6-$15',
    adj_value >= 16 & adj_value <= 25 ~ '$16-$25',
    adj_value >= 26 & adj_value <= 35 ~ '$26-$35',
    adj_value >= 36 & adj_value <= 50 ~ '$36-$50',
    adj_value >=51 & adj_value <= 75 ~ '$51-$75',
    adj_value > 75 ~ '$75+',
    TRUE ~ as.character(NA)
  )) %>% 
  mutate(top40 = ifelse(league_rank <= 40, 1, 0)) %>%
  group_by(year,bin) %>% 
  summarise(total = n(),
            total_top40 = sum(top40)) %>%
  mutate(pct_top40 = total_top40/total) %>% 
  ungroup() %>%
  filter(year >= 2010) 
ggplot(data = top40_by_year %>% mutate(fraction = paste0(total_top40,"/",total)),
       aes(x = bin, y = pct_top40, colour = as.factor(year))) +
  geom_text(aes(x = bin, y = pct_top40, label = fraction))

top40_by_year %>%
  group_by(bin) %>%
  summarise(mean_pct = mean(pct_top40))
```

## How much turnover YOY in top 5, 10, 20, 40?

```{r}
yoy_top <- 
wr %>%
  filter(is.na(league_rank) == FALSE) %>%
  mutate(top40 = ifelse(league_rank <= 40,1,0),
         top20 = ifelse(league_rank <= 20,1,0),
         top10 = ifelse(league_rank <= 10,1,0),
         top5  = ifelse(league_rank <= 5,1,0)) %>%
  arrange(player,year) %>%
  group_by(player) %>%
  mutate(top40_lastyear = ifelse(lag(year) == year - 1, lag(top40),NA),
         top20_lastyear = ifelse(lag(year) == year - 1, lag(top20),NA),
         top10_lastyear = ifelse(lag(year) == year - 1, lag(top10),NA),
         top5_lastyear  = ifelse(lag(year) == year - 1, lag(top5),NA),
         team_rank_lastyear = ifelse(lag(year) == year - 1,lag(team_rank),NA),
         league_rank_lastyear = ifelse(lag(year) == year - 1,lag(league_rank),NA))
```

```{r}
yoy_top %>%
  filter(top20 == 1) %>%
  group_by(year) %>%
  summarise(total_players = sum(top20),
            carryover = sum(top20_lastyear, na.rm = T))
```

```{r}
yoy_top %>%
  filter(top10 == 1) %>%
  group_by(year) %>%
  summarise(total_players = sum(top10),
            carryover10 = sum(top10_lastyear,na.rm = T),
            carryover20 = sum(top20_lastyear,na.rm = T),
            carrovery40 = sum(top40_lastyear,na.rm = T))
```

```{r}
yoy_top %>%
  filter(top5 == 1) %>%
  group_by(year) %>%
  summarise(total_players = sum(top5),
            carryover5 = sum(top5_lastyear, na.rm = T),
            carryover10 = sum(top10_lastyear,na.rm = T),
            carryover20 = sum(top20_lastyear,na.rm = T),
            carrovery40 = sum(top40_lastyear,na.rm = T))
```



## Can we predict top 20 year over year?

Create features such as lag in team rank, league rank


```{r}
head(yoy_top)
```
```{r}
library(xgboost)
model <- xgboost(data = as.matrix(yoy_top[,c(12:17)]), 
        label = yoy_top$top20,
        nrounds = 5,
        objective = "binary:logistic")
str(model)

preds <- predict(model, newdata = as.matrix(yoy_top[,c(12:17)]))
pred_df <- data.frame(year = yoy_top$year,
                      actual = yoy_top$top20,
                      pred = preds,
                      error = preds - yoy_top$top20) 
ggplot(data = pred_df,
       aes(x = pred, y = actual)) +
         geom_jitter() +
  geom_smooth()

xgb.importance(feature_names = colnames(yoy_top)[c(12:17)], model = model)
```

```{r}
model <- lm(data = pred_df,
            actual ~ pred)
ggplot(data = data.frame(r = model$resid,
                         year = pred_df$year),
       aes(x = year, y = r)) +
  geom_point() +
  geom_smooth()
```
```{r}
pROC::auc(pred_df$actual,pred_df$pred)
pROC::auc(yoy_top$top20,yoy_top$top20_lastyear)
```


## top 10 model
```{r}
library(xgboost)
top10_model <- xgboost(data = as.matrix(yoy_top[,c(12:17)]), 
        label = yoy_top$top10,
        nrounds = 5,
        objective = "binary:logistic")

pred_df <- data.frame(year = yoy_top$year,
                      actual = yoy_top$top10,
                      pred = predict(top10_model, newdata = as.matrix(yoy_top[,c(12:17)]))) 
ggplot(data = pred_df,
       aes(x = pred, y = actual)) +
         geom_jitter() +
  geom_smooth()

pROC::auc(pred_df$actual,pred_df$pred)

xgb.importance(feature_names = colnames(yoy_top)[c(12:17)], model = top10_model)

```

What if we just assumed the ten from last year would be the ten from this year?

```{r}
pROC::auc(yoy_top$top10,yoy_top$top10_lastyear)
```

## Build Iterative Models

* Incorporate age and experience
* Using 2000, predict 2001,
* Using 2000 and 2001, predict 2002, etc.

Can we use outputs from other models as inputs to the current?
So can we use P(X is top 5 in 2002) ~ P(X is top 10 in 2002) + P(X is top 20 in 2002) + P(X is top 40 in 2002)
This would be a model enhancement, so we could have a vanilla model built using only lag ranks and values then add in other inputs to enhance
the model, maybe it will work, maybe it won't.

Keep in mind the idea is to increase the probability we find a WR who will be in the top X at certain price points beyond what the raw data 
would say. 



### Look at season sd x average plots
```{r}
head(wr)
```

```{r}
wr_year <-
  wr %>%
  filter(league_rank <= 60) %>%
  inner_join(wr_offense, by = c('player','year')) %>% 
  group_by(player,name.x,year) %>%
  summarise(avg = mean(points),
            std = sd(points),
            games = n())
```


```{r}
ggplot(data = wr_year %>% filter(games >= 12),
       aes(x = avg, y = std)) +
  geom_point()
```

```{r}
?kmeans
wr_year_matrix <- wr_year %>% ungroup() %>% filter(games >= 12) %>% select(avg,std) %>% as.matrix()
k <- kmeans(wr_year_matrix, 7, nstart = 20)

ggplot(data = data.frame(k = k$cluster, avg = wr_year_matrix[,1], std = wr_year_matrix[,2]),
       aes(x = avg, y = std, colour = as.factor(k))
       ) +
  geom_point()
  

group5 <- which(k$cluster == 5)

wr_year_matrix2 <- wr_year %>% ungroup() %>% filter(games >= 12) 
wr_year_matrix3 <- wr_year_matrix2[group5,]
group5_draft <- left_join(wr_year_matrix3, draft_data, by = c('player' = 'player_code','year')) %>% filter(year >= 2010)
head(group5_draft)
```
